---
title: "Homework 2"
author: "Jackson Connor"
format: html
editor: visual
---

## Packages

```{r}
#| warning: false
library(tidyverse)
library(multcomp)
library(emmeans)
```

## Problem 1

A pharmaceutical manufacturer wants to investigate the bioactivity of a new drug. A completely randomized single-factor experiment was conducted with three dosage levels, and the following results were obtained.

Complete the ANOVA Table: ANOVA Table

|            | df  | SS    | MS  | F-value | P-value |
|------------|-----|-------|-----|---------|---------|
| Bet_Dosage |     | 450.7 |     |         |         |
| Error      |     |       |     |         |         |
| Total      |     | 738.9 |     |         |         |

```         
Is there evidence to indicate that dosage level affects bioactivity? 

Compute a 95 percent interval estimate of the mean of dosage level is 30g. 

Compute a 99 percent interval estimate of the mean difference between dosages levels 20g and 40g.

Compute simultaneous 95% CIs for the difference in means for dosage levels 20g and 40g, based on Tukey’s, Bonferroni’s and Scheffe’s Methods:
```

Note: Q=q\_((1-α), a,(n_T-a))/√2=q\_(0.95, 4,(3\*4-3))/√2=3.95/√2=2.79

```         
Use Tukey’s HSD to compare pairs of treatment means. 
```

## Problem 2

The tensile strength of Portland cement is being studied. Four different mixing techniques can be used economically. A completely randomized experiment was conducted, and the following data were collected.

You are given the following information:

Y ̅*(1.)=2971.00,s(Y ̅*(1.) )=56.65 and MSE=12825.7

Y ̅*(2.)=3156.25,s(Y ̅*(2.) )=56.65

Y ̅*(3.)=2933.75,s(Y ̅*(3.) )=56.65

Y ̅*(4.)=2666.25,s(Y ̅*(4.) )=56.65

**a.** Test the hypothesis that mixing techniques affect the strength of the cement.

```{r}
Tensile_Strength=c(3129,3000,2865,2890,3200,3300,2975,3150,2800,2900,2985,3050,2600,2700,2600,2765)

Mixing_Technique=c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,4)

data<-data.frame(Tensile_Strength,Mixing_Technique)

str(data)

Mixing_Technique = factor(Mixing_Technique)

model <- aov(Tensile_Strength ~ Mixing_Technique)

summary(model)
```


**b.** Find a 95 percent confidence interval on the mean tensile strength of the Portland cement produced by mixing technique 3.

$\bar{y_3}$ = 2933.7, s($\bar{y_3}$) = 56.65

```{r}
qt(0.975, 12)
```


$2933.7 \pm (2.179*56.65)$ = **(2810.26, 3057.14)**

**c.** Find a 95 percent confidence interval on the difference in means for techniques 1 and 3 & 3 and 4.

$L = \frac{\mu_1 + \mu_3}{2}-\frac{\mu_3+\mu_4}{2} = (\frac{1}{2})\mu_1 + (0)\mu_2 + (\frac{1}{2})(\frac{-1}{2})\mu_3 + (\frac{-1}{2})\mu_4$

$\hat{L} = \frac{\hat{y_1}+\hat{y_3}}{2}-\frac{\hat{y_3}+\hat{y_4}}{2} = \frac{{2971}+{2933.75}}{2}-\frac{{2933.75}+{2666.25}}{2}$ = 152.375

$\overset{4}{\sum}_{i=1}\frac{c^2_i}{n_i} = {\frac{\frac{1}{2}^2}{4}} + {\frac{\frac{-1}{4}^2}{4}} + {\frac{\frac{-1}{2}^2}{4}}$ = 

**d.** Compute simultaneous 95% CIs for the difference in means for techniques 1 and 3, based on Tukey’s, Bonferroni’s and Scheffe’s Methods:

```{r}
modelemm <- emmeans(model, "Mixing_Technique")

pairs(modelemm, infer=c(T,T), level=0.95, adjust="tukey")

```
(-200.5, 275.0)

```{r}
modelemm <- emmeans(model, "Mixing_Technique")

pairs(modelemm, infer=c(T,T), level=0.95, adjust="bonferroni")

```
(-215.2, 289.7)

```{r}

modelemm <- emmeans(model, "Mixing_Technique")

pairs(modelemm, infer=c(T,T), level=0.95, adjust="scheffe")
```
(-221.88, 296.4)

**e.** Compute Bonferroni’s Minimum Significant Difference

$t_{(1-\frac{\alpha}{2*g}), df_{error}}$ = Bonferroni's CV

g = $\frac{a(a-1)}{2} = \frac{4(3)}{2}$ = 6
$1-\frac{0.05}{2*6}$ = 1 - .0042 = .9958


```{r}
qt(.9958, 12)
```
$B_{CV} * SE$ = 3.14839 * 80.1 = 252.186

**f.** Compute Tukey’s Honest Significant Difference

$qtukey_{1-\alpha, df_{btwn}, df_{error}}$ = Tukey's CV

```{r}
qtukey(0.95, 3, 12)
```
$3.772929 \div \sqrt{2}$ = 2.668

$Tukey_{CV} * SE$ = 2.668 * 80.1 = 213.696

**g.** Compute Scheffe’s Minimum Significant Difference

$\sqrt{(a-1)f_{(1-\alpha),(a-1),df_{error}}}$ = Scheffe's CV

```{r}
qf(0.95,3,12)
```
S = $\sqrt{3*3.490295}$ = 3.236

$S_{CV} * SE$ 3.236 * 80.1 = 259.194
