---
title: "HW3"
author: "Jackson Connor"
format: html
editor: visual
---

## Problem 1

A regression analysis relating test scores (Y) to training hours (X) produced the following fitted equation: y ̂=25-0.5x.

**a.** What is the fitted value of the response variable corresponding to x = 7?

$\hat{y} = 25-0.5(7) = 21.5$

**b.** What is the residual corresponding to the data point with x = 3 and y = 30? Is the point above or below the line? Why?

$\hat{y} = 25-0.5x = 25-0.5(3) = 23.5$

residual = $e_i = y_i - \hat{y}_i = 30-23.5 = 6.5$

The residual corresponding to the data point is above the line since the calculated residual is a positive value.

**c.** If x increases 3 units, how does y ̂ change?

For each one unit increase of x, y decreases by 0.5; so, for a 3 unit increase of x, y would decrease 1.5 units.

**d.** An additional test score is to be obtained for a new observation at x = 6. Would the test score for the new observation necessarily be 22? Explain.

The test score would not necessarily be 22 as the model is used to estimate values and built on a sample set of observations. If you used the model to estimate a value, it would be 22, however the actual value of the observation may differ slightly.

**e.** The error sums of squares (SSE) for this model were found to be 7. If there were n = 16 observations, provide the best estimate for σ\^2.

$\hat{\sigma}^2 = MSE = \frac{SSE}{(n-p)} = \frac{7}{16-1} = \frac{7}{15} = 0.467$

## Problem 2

The dataset "Healthy Breakfast" contains, among other variables, the Consumer Reports ratings of 77 cereals and the number of grams of sugar contained in each serving. Considering "Sugars" as the explanatory variable and "Rating" as the response variable generated the following fitted regression equation:

(rating) ̂=59.3-2.40\*sugars

The "Analysis of Variance" partial portion of the R output is shown below.

| Source     | DF     | SS         | MS         | F           | P            |
|------------|--------|------------|------------|-------------|--------------|
| Regression | **1**  | 8654.7     | 8654.7     | **102.348** | **1.11e-15** |
| Error      | **75** | **6342.1** | **84.561** |             |              |
| Total      | **76** | 14996.8    |            |             |              |

**a.** Find the missing values.

$df_{reg} = 2-1 = 1$

$df_{error} = 77-2 = 75$

$df-{total} = 75+1 = 76$

SSE = SST - SSR = 14996.8 - 8654.7 = 6342.1

MSE = $\frac{SSE}{n-p} = \frac{6342.1}{75} = 84.561$

F = $\frac{MSR}{MSE} = \frac{8654.7}{84.561} = 102.348$

```{r}
1-pf(102.348, 1, 75)
```

**b\.** Find $R^2$ value and interpret that number.

$R^2 = \frac{SSR}{SST} = \frac{8654.7}{14996.8} = 0.577$

57.7% of variation in the ratings is due to the sugars in the cereal.

**c\.** What is the estimated value of $\sigma^2$?

$\sigma^2 = MSE = 84.561$

**d\.** Test $H_0: \beta_1 = 0$ against $H_a:\beta_1 \ne 0$ using F-test

1.11e-15 \< 0.05, therefore we reject the null hypothesis and have strong evidence to conclude the alternative that $\beta_1 \neq 0$, ie that there is a strong linear relationship.

## Problem 3
Athletes are constantly seeking measures of the degree of their cardiovascular fitness prior to a major race. Athletes want to know when their training is at a level that will produce a peak performance. One such measure of fitness is the time to exhaustion from running on a treadmill at a specified angle and speed. The important question is then “Does this measure of cardiovascular fitness translate into performance in a 10-km running race?” Twenty experienced distance runners who professed to be in top condition were evaluated on the treadmill and then
had their times recorded in a 10-km race.


You are provided with the following information:

β^_0=b_0=58.816,β^_1=b_1=-1.867,s(b_1 )=0.346,MSE=4.417


**a.** Test the hypothesis that there is a linear relationship between the amount of time needed to run a 10-km race and the time to exhaustion on a treadmill.

$H_0: \beta_1 = 0$

$H_a: \beta_1 \neq 0$

$t* = \frac{b_1-0}{s{b_1}} = \frac{-1.867}{0.346} = -5.396$

```{r}
2*(1-pt(5.396, 18))
```
3.972e-05 < 0.05, so we reject our null that $\beta_1$ is equal to 0 and have strong evidence to conclude there is a linear relationship between the amount of time needed to run a 10km race and the time to exhaustion on a treadmill.

**b.** Construct a 95% confidence interval for β1.

$\hat{\beta}_1 \pm t_{1-\frac{\alpha}{2},df_{error}}*sd(\hat{\beta}_1)$ = -1.867 $\pm$ (2.101*0.346)

```{r}
qt(.975, 18)
```
(-2.59, -1.14)

**c.** Construct a 95% confidence interval for the estimated mean 10-km race when the treadmill time is 10 minutes. Given that sd{y ̂_(x_h=10)}=0.7342

Point Estimate: $y(10) = 58.816 + (-)1.867(10) = 40.146

40.146 $\pm$ 2.101(0.7324)

(38.60, 41.69)


**d.** Construct a 95% prediction interval for an athlete whose treadmill time is 10 minutes. sd(y ̂_pred )=√(mse+(sd{y ̂_(x_h=10)})^2 )=2.226

40.146 $\pm$ 2.101(2.226)

(35.47, 44.82)
